---
title: "Bellabeat Case Study - Processing File"
author: "Kevin Leung"
date: "2025-01-27"
output: html_document
---

## Setting up my environment

Setting up my R environment by loading 'tidyverse', 'here', 'skimr' and 'janitor' packages

```{r loading packages}
install.packages("tidyverse")
install.packages("here")
install.packages("skimr")
install.packages("janitor")
library("tidyverse")
library("here")
library("skimr")
library("janitor")
```

## Normalization of Data

Loading, previewing and standardizing the dataset files

**Reading dataset files**

```{r reading dataset files}
daily_activity_mar_apr <- read_csv("dailyactivity_3_to_4.csv")
heart_rate_mar_apr <- read_csv("heartrate_seconds_3_to_4.csv")
hourly_calories_mar_apr <- read_csv("hourlycalories_3_to_4.csv")
hourly_intensities_mar_apr <- read_csv("hourlyintensities_3_to_4.csv")
hourly_steps_mar_apr <- read_csv("hourlysteps_3_to_4.csv")
minute_calories_mar_apr <- read_csv("minutecaloriesnarrow_3_to_4.csv")
minute_intensities_mar_apr <-read_csv("minuteintensitiesnarrow_3_to_4.csv")
minute_mets_mar_apr <- read_csv("minutemetsnarrow_3_to_4.csv")
minute_sleep_mar_apr <- read_csv("minutesleep_3_to_4.csv")
minute_steps_mar_apr <- read_csv("minutestepsnarrow_3_to_4.csv")
weight_log_mar_apr <- read_csv("weightloginfo_3_to_4.csv")
daily_activity_apr_may <- read_csv("dailyactivity_4_to_5.csv")
daily_calories_apr_may <- read_csv("dailycalories_4_to_5.csv")
daily_intensities_apr_may <- read_csv("dailyintensities_4_to_5.csv")
daily_steps_apr_may <- read_csv("dailysteps_4_to_5.csv")
heart_rate_apr_may <- read_csv("heartrate_seconds_4_to_5.csv")
hourly_calories_apr_may <- read_csv("hourlycalories_4_to_5.csv")
hourly_intensities_apr_may <- read_csv("hourlyintensities_4_to_5.csv")
hourly_steps_apr_may <- read_csv("hourlysteps_4_to_5.csv")
minute_calories_apr_may <- read_csv("minutecaloriesnarrow_4_to_5.csv")
minute_calories_apr_may_2 <- read_csv("minutecalorieswide_4_to_5.csv")
minute_intensities_apr_may <- read_csv("minuteintensitiesnarrow_4_to_5.csv")
minute_intensities_apr_may_2 <- read_csv("minuteintensitieswide_4_to_5.csv")
minute_mets_apr_may <- read_csv("minutemetsnarrow_4_to_5.csv")
minute_sleep_apr_may <- read_csv("minutesleep_4_to_5.csv")
minute_steps_apr_may <- read_csv("minutestepsnarrow_4_to_5.csv")
minute_steps_apr_may_2 <- read_csv("minutestepswide_4_to_5.csv")
sleep_day_apr_may <- read_csv("sleepday_4_to_5.csv")
weight_log_apr_may <- read_csv("weightloginfo_4_to_5.csv")
```

**Previewing data for n_missing values**

```{r previewing data}
skim_without_charts(daily_activity_mar_apr)
head(daily_activity_mar_apr)
skim_without_charts(heart_rate_mar_apr)
head(heart_rate_mar_apr)
skim_without_charts(hourly_calories_mar_apr)
head(hourly_calories_mar_apr)
skim_without_charts(hourly_intensities_mar_apr)
head(hourly_intensities_mar_apr)
skim_without_charts(hourly_steps_mar_apr)
head(hourly_steps_mar_apr)
skim_without_charts(minute_calories_mar_apr)
head(minute_calories_mar_apr)
skim_without_charts(minute_intensities_mar_apr)
head(minute_intensities_mar_apr)
skim_without_charts(minute_mets_mar_apr)
head(minute_mets_mar_apr)
skim_without_charts(minute_sleep_mar_apr)
head(minute_sleep_mar_apr)
skim_without_charts(minute_steps_mar_apr)
head(minute_steps_mar_apr)
skim_without_charts(weight_log_mar_apr)
head(weight_log_mar_apr)
skim_without_charts(daily_activity_apr_may)
head(daily_activity_apr_may)
skim_without_charts(daily_calories_apr_may)
head(daily_calories_apr_may)
skim_without_charts(daily_intensities_apr_may)
head(daily_intensities_apr_may)
skim_without_charts(daily_steps_apr_may)
head(daily_steps_apr_may)
skim_without_charts(heart_rate_apr_may)
head(heart_rate_apr_may)
skim_without_charts(hourly_calories_apr_may)
head(hourly_calories_apr_may)
skim_without_charts(hourly_intensities_apr_may)
head(hourly_intensities_apr_may)
skim_without_charts(hourly_steps_apr_may)
head(hourly_steps_apr_may)
skim_without_charts(minute_calories_apr_may)
head(minute_calories_apr_may)
skim_without_charts(minute_calories_apr_may_2)
head(minute_calories_apr_may_2)
skim_without_charts(minute_intensities_apr_may)
head(minute_intensities_apr_may)
skim_without_charts(minute_intensities_apr_may_2)
head(minute_intensities_apr_may_2)
skim_without_charts(minute_mets_apr_may)
head(minute_mets_apr_may)
skim_without_charts(minute_sleep_apr_may)
head(minute_sleep_apr_may)
skim_without_charts(minute_steps_apr_may)
head(minute_steps_apr_may)
skim_without_charts(minute_steps_apr_may_2)
head(minute_steps_apr_may_2)
skim_without_charts(sleep_day_apr_may)
head(sleep_day_apr_may)
skim_without_charts(weight_log_apr_may)
head(weight_log_apr_may)
```

**Previewing column names to determine a relationship**

```{r previewing column relationships}
colnames(daily_activity_mar_apr)
colnames(heart_rate_mar_apr)
colnames(hourly_calories_mar_apr)
colnames(hourly_intensities_mar_apr)
colnames(hourly_steps_mar_apr)
colnames(minute_calories_mar_apr)
colnames(minute_intensities_mar_apr)
colnames(minute_mets_mar_apr)
colnames(minute_sleep_mar_apr)
colnames(minute_steps_mar_apr)
colnames(weight_log_mar_apr)
colnames(daily_activity_apr_may)
colnames(daily_calories_apr_may)
colnames(daily_intensities_apr_may)
colnames(daily_steps_apr_may)
colnames(heart_rate_apr_may)
colnames(hourly_calories_apr_may)
colnames(hourly_intensities_apr_may)
colnames(hourly_steps_apr_may)
colnames(minute_calories_apr_may)
colnames(minute_calories_apr_may_2)
colnames(minute_intensities_apr_may)
colnames(minute_intensities_apr_may_2)
colnames(minute_mets_apr_may)
colnames(minute_sleep_apr_may)
colnames(minute_steps_apr_may)
colnames(minute_steps_apr_may_2)
colnames(sleep_day_apr_may)
colnames(weight_log_apr_may)
```

**Identifying number of unique ids**

```{r identifying unique ids}
daily_activity_id <- n_distinct(daily_activity_mar_apr$Id)
heart_rate_id <- n_distinct(heart_rate_mar_apr$Id)
hourly_calories_id <- n_distinct(hourly_calories_mar_apr$Id)
hourly_intensities_id <- n_distinct(hourly_intensities_mar_apr$Id)
hourly_steps_id <- n_distinct(hourly_steps_mar_apr$Id)
minute_calories_id <- n_distinct(minute_calories_mar_apr$Id)
minute_intensities_id <- n_distinct(minute_intensities_mar_apr$Id)
minute_mets_id <- n_distinct(minute_mets_mar_apr$Id)
minute_sleep_id <- n_distinct(minute_sleep_mar_apr$Id)
minute_steps_id <- n_distinct(minute_steps_mar_apr$Id)
weight_log_id <- n_distinct(weight_log_mar_apr$Id)
daily_activity_id2 <- n_distinct(daily_activity_apr_may$Id)
daily_calories_id2 <- n_distinct(daily_calories_apr_may$Id)
daily_intensities_id2 <- n_distinct(daily_intensities_apr_may$Id)
daily_steps_id2 <- n_distinct(daily_steps_apr_may$Id)
heart_rate_id2 <- n_distinct(heart_rate_apr_may$Id)
hourly_calories_id2 <- n_distinct(hourly_calories_apr_may$Id)
hourly_intensities_id2 <- n_distinct(hourly_intensities_apr_may$Id)
hourly_steps_id2 <- n_distinct(hourly_steps_apr_may$Id)
minute_calories_id2 <- n_distinct(minute_calories_apr_may$Id)
minute_calories_id3 <- n_distinct(minute_calories_apr_may_2$Id)
minute_intensities_id2 <- n_distinct(minute_intensities_apr_may$Id)
minute_intensities_id3 <- n_distinct(minute_intensities_apr_may_2$Id)
minute_mets_id2 <- n_distinct(minute_mets_apr_may$Id)
minute_sleep_id2 <- n_distinct(minute_sleep_apr_may$Id)
minute_steps_id2 <- n_distinct(minute_steps_apr_may$Id)
minute_steps_id3 <- n_distinct(minute_steps_apr_may_2$Id)
sleep_day_id2 <- n_distinct(sleep_day_apr_may$Id)
weight_log_id2 <- n_distinct(weight_log_apr_may$Id)
```

**Identifying number of total observations**

```{r identifying number of observations}
daily_activity_mar_apr %>% nrow()
heart_rate_mar_apr %>% nrow()
hourly_calories_mar_apr %>% nrow()
hourly_intensities_mar_apr%>% nrow()
hourly_steps_mar_apr %>% nrow()
minute_calories_mar_apr %>% nrow()
minute_intensities_mar_apr %>% nrow()
minute_mets_mar_apr %>% nrow()
minute_sleep_mar_apr %>% nrow()
minute_steps_mar_apr %>% nrow()
weight_log_mar_apr %>% nrow()
daily_activity_apr_may %>% nrow()
daily_calories_apr_may %>% nrow()
daily_intensities_apr_may %>% nrow()
daily_steps_apr_may %>% nrow()
heart_rate_apr_may %>% nrow()
hourly_calories_apr_may %>% nrow()
hourly_intensities_apr_may %>% nrow()
hourly_steps_apr_may %>% nrow()
minute_calories_apr_may %>% nrow()
minute_calories_apr_may_2 %>% nrow()
minute_intensities_apr_may %>% nrow()
minute_intensities_apr_may_2 %>% nrow()
minute_mets_apr_may %>% nrow()
minute_sleep_apr_may %>% nrow()
minute_steps_apr_may %>% nrow()
minute_steps_apr_may_2 %>% nrow()
sleep_day_apr_may %>% nrow()
weight_log_apr_may %>% nrow()
```
